On the specific recursive structure of iterated parametricity

### 2025-09-28 paper/paper.tex
Moving a paragraph to prevent a big overflow (even if still a small underflow)
paper: start new paper

\documentclass{art.cls/art}

\usepackage{environ, multirow, tabularx, caption, booktabs, makecell, colortbl, minted, art.cls/joinargs}
\setminted[]{fontfamily=lmtt}
\usepackage[prefix=bonak]{art.cls/xkeymask}

@@

% Simplicial and Cubical sets
\newcommand{\DeltaHat}{\ensuremath{\hat{\boldsymbol{\Delta}}}}
\newcommand{\DeltaPlus}{\ensuremath{\boldsymbol{\Delta}_+}}
\newcommand{\SSSet}{\ensuremath{\Set_{\DeltaPlus}}}
\newcommand{\Cube}{\ensuremath{\boldsymbol{{\square}}}}
\newcommand{\CSet}{\ensuremath{\Set_{\Cube}}}

% Universe, referred to as U
\renewcommandx{\U}[1][1=]{\ensuremath{\mathsf{\color{spanish-blue}{HSet}}_{#1}}}
\newcommandx{\HGpd}[1][1=]{\ensuremath{\mathsf{\color{spanish-blue}{HGpd}}_{#1}}}
\newcommand{\Type}{\ensuremath{\mathsf{\color{spanish-blue}{Type}}}}
\newcommand{\SProp}{\ensuremath{\mathsf{\color{spanish-blue}{SProp}}}}

@@

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\newcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\newcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\newcommandx{\restrc}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\newcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\newcommandx{\cohl}{\coh{layer}[][][][][][][][]}
\newcommandx{\cohc}{\coh{painting}[][][][][][][][]}
\newcommandx{\coht}{\cohtwo{frame}[][][][][][][][]}
\newcommandx{\fullframe}[1][1]{\prim{fullframe}[][#1][][][]}

@@

% Caption setup
\DeclareCaptionFormat{plain}{#1: #3}
\captionsetup{font=footnotesize,labelfont=bf}

% The eqntable environment
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\NewEnviron{eqntable}[1]{
  \begin{table}[h]
  \small
    \begin{tabularx}{0.94\linewidth}{
      @{}
      >{$}l<{$}
      >{$}c<{$}
      >{$}c<{$}
      >{$}Y<{$}
      @{}}
      \toprule
      \BODY
      \bottomrule
    \end{tabularx}
    \caption{#1}
  \end{table}
}

@@

% Set the entire document to gray65, and pick certain colors
\color{gray65}

\begin{document}
\title{A parametricity-based formalization of semi-simplicial and semi-cubical sets}
\date{}
\author{
  \textcolor{gray80}{Hugo Herbelin} \\
  \itshape \textcolor{gray80}{Université Paris-Cité, Inria, CNRS, IRIF, Paris} \\
  \ttfamily \textcolor{gray80}{Hugo.Herbelin@inria.fr}
  \and
  \textcolor{gray80}{Ramkumar Ramachandra} \\
  \itshape \textcolor{gray80}{Université Paris-Cité (2020-2022), Unaffiliated} \\
  \ttfamily \textcolor{gray80}{r@artagnon.com}
}
\maketitle
\begin{abstract}
  Semi-simplicial and semi-cubical sets are commonly defined as presheaves over respectively, the semi-simplex or semi-cube category. Homotopy Type Theory then popularized an alternative definition, where the set of $n$-simplices or $n$-cubes are instead regrouped into the families of the fibres over their faces, leading to a characterization we call \emph{indexed}. Moreover, it is known that semi-simplicial and semi-cubical sets are related to iterated Reynolds parametricity, respectively in its unary and binary variants. We exploit this correspondence to develop an original uniform indexed definition of both augmented semi-simplicial and semi-cubical sets, and fully formalize it in Coq.
\end{abstract}

\section{Introduction}
\subsection*{Fibred vs indexed presentation of semi-simplicial and semi-cubical sets}
A family of sets can commonly be represented in two ways: as a family properly speaking, indexed by the elements of a given set $S$, or as a set $T$ together with a map from $T$ to $S$, which specifies for each element of $T$ its dependency on $S$. In the former case, we call it an \emph{indexed} presentation. In the latter case, the set associated to a given element of $S$ is the fibre of this element, so we call it a \emph{fibred} presentation. The two presentations are equivalent and the equivalence can be phrased concisely in the language of homotopy type theory~\cite{hottbook} as the fibred/indexed equivalence\footnote{In an informal discussion, alternative nomenclatures were proposed: fibration/family equivalence and unbundled/bundled equivalence. The fibred/indexed nomenclature echoes the Grothendieck construction of fibred categories from indexed categories. The most elementary instance of the equivalence, with $\Type$ instead of $\U$, is sometimes called ``Grothendieck construction for dummies'', and its proof requires univalence~\cite{hottbook}.}.
\begin{equation*}
  \mbox{(fibred)}\qquad(\Sigma T: \U. (T \rightarrow S)) ~\simeq~ (S \rightarrow \U)\qquad \mbox{(indexed)}
\end{equation*}
Here, $\U$ represents in homotopy type theory the subset of types within a given universe where equality of any two elements has at most one proof.

A \emph{presheaf} on an category is a family of sets indexed by the object of the category with maps indexed by the morphisms. As such, it lives on the indexed side of the equivalence, contrasting with the fibred side, where we have \emph{discrete Grothendieck fibrations}~\cite{LoregianRiehl20}. However, there are situations where a presheaf can also be seen as living on the fibred side of the equivalence. This happens when the indexing category is \emph{direct}, or has a downwards-well-founded collection of non-identity morphisms. Let us consider, for instance, the case of a semi-cubical set~\cite{grandis03,buchholtz17} presented with $2n$ face maps from the set of $n$-cubes to the set of $(n-1)$-cubes. Formulated in type theory, the corresponding presheaf definition of a semi-cubical set prescribes a family of sets and face maps between them as follows.

@@

\subsection*{Reynolds parametricity and its unary and binary variants}
In the context of functional programming, Reynolds parametricity~\cite{reynolds83} interprets types as relations characterizing the observational behaviour of programs of this type. More generally, families over a product of sets, or \emph{correspondences}, can be used in place of relations. Parametricity can then be iterated, and relying on the fibred presentation of correspondences as spans, it has been noted that iterated Reynolds parametricity has the structure of a cubical set~\cite{altenkirch15,moulin16,johann17,moeneclaey21,moeneclaey22phd}. We obtain a \emph{unary} variant of Reynolds \emph{binary} parametricity by using predicates or families instead of relations or correspondences, and this is a form of realizability~\cite{bernardy12,lasson12,moulin16}. Cubical set models which differ only by the arity one~\cite{bernardy15} or two~\cite{bezem13} were introduced, and this led to a general notion of affine $\nu$-ary cubes in \cite{nuytsdevriese24}. In parallel, it has been noted that iterated unary parametricity has the structure of an augmented simplicial set\footnote{Private communication with Hugo Moeneclaey and Thorsten Altenkirch.}. This suggests that the definition of augmented semi-simplicial sets and semi-cubical sets can in turn be seen as particular instances of the restriction of Nuyts-Devriese's $\nu$-ary cubes to only faces, which we call $\nu$-sets, of presheaves over a $\nu$-semi-shape category made of words of some cardinal $\nu+1$, where $\nu=1$ gives augmented semi-simplicial sets and $\nu=2$ gives semi-cubical sets.

\subsection*{Contribution}
The main contribution of the paper is to describe the details of a recipe that uniformly characterizes unary and binary iterated parametricity in indexed form, and to derive from it a new indexed presentation, called indexed \emph{$\nu$-sets}, of augmented semi-simplicial and semi-cubical sets.

Our work is a step in the direction of the program initiated in \cite{altenkirch15} to develop parametricity-based models of parametric type theory~\cite{bernardy15,nuyts17,cavallo19} and cubical type theory~\cite{bezem13,cohen16,angiuli21}, which are closer to the syntax of type theory, and are likely to better reflect the definitional properties of type theory than presheaf-based cubical sets would. For example, consider the loss of definitional properties when interpreting ``indexed'' dependent types of type theory as ``fibrations'' in models such as locally cartesian closed categories~\cite{curien14}.

Our construction has the unique property of reflecting the structure of parametricity and of yielding both augmented semi-simplicial and semi-cubical sets from the same construction. The approach taken in \cite{part15} and \cite{altenkirch16} takes benefit of the definitional compositionality of increasing injective maps, but we do not see how they could be generalized to yield semi-cubical sets.

Our mechanization can be found at \href{https://github.com/artagnon/bonak}{github.com/artagnon/bonak}. The construction was conceived in Summer 2019, and the mechanization began in late 2019. A sketch of the construction was presented at the 2020 HoTT-UF workshop, and the completion of the mechanization was reported at the TYPES 2022 conference.

\section{Semi-simplicial and semi-cubical sets\label{sec:nu}}
In this section, we generalize semi-simplicial and semi-cubical sets to $\nu$-sets, subsuming the earlier definitions. We start with some introductory material on semi-simplicial and semi-cubical sets.

\subsection*{Augmented semi-simplicial sets}
Augmented semi-simplicial sets are defined similarly to semi-simplicial sets, except that the connected components are additionally dependent on a ``colour''. Conversely, semi-simplicial sets can be seen as augmented semi-simplicial sets over the singleton set of a fixed colour. Let us associate dimension $-1$ to colours; then, points are dimension $0$, lines are dimension $1$, and so on.

Ordinary semi-simplicial sets are presheaves over the semi-simplex category. Augmented semi-simplicial sets are presheaves over \DeltaPlus. There are different ways to define \DeltaPlus, up to equivalence, and we use a definition that can be extended to semi-cubical sets in a straightforward manner. In particular, we start numbering objects from $0$ instead of $-1$ so that there is a shift by one compared to the standard numbering of augmented semi-simplicial sets.

\begin{notation}[Finite sequences]
  We denote finite sequences by $i_1 \ldots i_n$ for $i_j$ ranging over some domain. The empty sequence is written $\epsilon$.
\end{notation}

\begin{definition}[$\DeltaPlus$]
  The definition of $\DeltaPlus$ is shown below. Note that, if $g \circ f$ is well-defined, then the length of $f$ is less than or equal to that of $g$. It can be shown that composition is associative and that $\id$ is neutral.
  \begin{align*}
    \Obj_{\DeltaPlus}       & := \Nat                                                                         \\
    \Hom_{\DeltaPlus}(p, n) & := \{l \in \{0, \kstar\}^n \mid \text{number of $\kstar$ in $l = p$}\}          \\
    g \circ f               & :=
    \begin{cases}
      f                & \text{if $g = \epsilon$}                                           \\
      0\,(g' \circ f)  & \text{if $g = 0\,g'$}                                              \\
      a\,(g' \circ f') & \text{if $g = \kstar\,g'$, $f = a\,f'$, where $a = 0$ or $\kstar$} \\
    \end{cases}                     \\
    \id                     & := \kstar \ldots \kstar \text{ $n$ times for $\id \in \Hom_{\DeltaPlus}(n, n)$}
  \end{align*}
\end{definition}

\begin{definition}[$\SSSet$]
  We define the category of augmented semi-simplicial sets as the functor category:
  \begin{equation*}
    \SSSet := \Set^{\DeltaPlus^{\mathsf{op}}}
  \end{equation*}
\end{definition}

To provide examples, we define the standard augmented $n$-semi-simplex, taking into account the shift by one in the numbering.

\begin{definition}[$\DeltaPlus^n$]
  The standard augmented $(n-1)$-semi-simplex $\DeltaPlus^{n-1}$ is defined as the Yoneda embedding of $n \in \Obj(\DeltaPlus)$:
  \begin{align*}
    \DeltaPlus^{n-1}    & : \SSSet                   \\
    \DeltaPlus^{n-1}(p) & := \Hom(p, n)              \\
    \DeltaPlus^{n-1}(f) & := \lambda g .\, g \circ f
  \end{align*}
\end{definition}

The standard augmented $(-1)$-semi-simplex is a singleton made of one colour (in this case, black). Standard augmented $n$-semi-simplices for $n \geq 0$ have a geometric interpretation, and we illustrate them for dimensions $0$, $1$, and $2$.

\begin{example}[$\DeltaPlus^0$]
  The standard augmented $0$-semi-simplex can be pictured as a point, coloured black, corresponding to the unique morphism in $\Hom(0, 1)$. This point is the identity in $\Hom(1, 1)$; it is hence shown as a singleton $\kstar$.
  \begin{equation*}
    \begin{tikzcd}
      \kstar
    \end{tikzcd}
  \end{equation*}
\end{example}

\begin{example}[$\DeltaPlus^1$]
  The standard augmented $1$-semi-simplex is drawn as two points, given by $\Hom(1, 2)$, along with a line connecting them, given by $\Hom(2, 2)$. We use black to denote the unique morphisms in $\Hom(0, 1)$ and $\Hom(0, 2)$.
  \begin{equation*}
    \begin{tikzcd}
      \kstar0 \arrow[r, dash, "\kstar\kstar"] & 0\kstar
    \end{tikzcd}
  \end{equation*}
\end{example}

\begin{example}[$\DeltaPlus^2$]
  $\DeltaPlus^2$ is drawn as three points, given by $\Hom(1, 3)$, three lines connecting them, given by $\Hom(2, 3)$, and a triangular filler given by $\Hom(3, 3)$.
  \begin{equation*}
    \begin{tikzcd}
      & |[alias=F]|00\kstar \arrow[ddr, dash, "0\kstar\kstar"] & \\\\
      \kstar00 \arrow[rr, dash, "\kstar\kstar0"{name=T, below}]\arrow[uur, dash, "\kstar0\kstar"] && 0\kstar0 \\
      \arrow[rightarrow, from=F, to=T, phantom, "\kstar\kstar\kstar" description]
    \end{tikzcd}
  \end{equation*}
\end{example}

More generally, the standard augmented $(n + 1)$-semi-simplex can be obtained by taking a copy of the standard augmented $n$-semi-simplex serving as a base, and gluing on top of it another copy lifted by one dimension. In the second copy, the colour becomes an extra point, the points become lines connecting the points of the base to the extra point, and so on. In particular, the components of the base are those of the standard augmented $n$-semi-simplex postfixed by $0$ while the components of the lifted copy are postfixed by $\kstar$. Note that the components may be oriented by letting each $n$-dimensional component point to the $(n-1)$-dimensional component obtained by replacing the leftmost $\kstar$ of the $n$-dimensional component with $0$.

\subsection*{Semi-cubical sets}
Semi-cubical sets are defined like augmented semi-simplicial sets except that $\DeltaPlus$ is replaced by $\Cube$ in which we take sequences of $L$, $R$ and $\kstar$, instead of sequences of $0$ and $\kstar$.

\begin{definition}[\Cube]
  The definition of $\Cube$ is shown below. The symbols $L$ and $R$ indicate opposite faces of a cube.

  \begin{align*}
    \Obj_{\Cube}       & := \Nat                                                                   \\
    \Hom_{\Cube}(p, n) & := \{l \in \{L, R, \kstar\}^n \mid \text{number of $\kstar$ in $l = p$}\} \\
    g \circ f          & :=
    \begin{cases}
      f                & \text{if $g = \epsilon$}                                                 \\
      a\,(g' \circ f)  & \text{if $g = a\,g'$}, \text{where $a = L$ or $R$}                       \\
      a\,(g' \circ f') & \text{if $g = \kstar\,g'$, $f = a\,f'$, where $a = L$, $R$, or $\kstar$} \\
    \end{cases}    \\
    \id                & := \kstar \ldots \kstar \text{ $n$ times}
  \end{align*}

  Again, if $g \circ f$ is well-defined, then the length of $f$ is less than or equal to that of $g$. It can be shown that composition is associative and that $\id$ is neutral.
\end{definition}

\begin{definition}[\CSet]
  We define the category of semi-cubical sets as the functor category:
  \begin{equation*}
    \CSet := \Set^{\op{\Cube}}
  \end{equation*}
\end{definition}

\begin{definition}[$\Cube^n$]
  The standard semi-cube $\Cube^n$ is defined as the Yoneda embedding of $n \in \Obj(\Cube)$:
  \begin{align*}
    \Cube^n    & : \CSet                    \\
    \Cube^n(p) & := \Hom(p, n)              \\
    \Cube^n(f) & := \lambda g .\, g \circ f
  \end{align*}
\end{definition}

Standard $n$-semi-cubes have a geometric interpretation, which we illustrate for dimensions $0$, $1$, and $2$.

\begin{example}[$\Cube^0$]
  $\Cube^0$ is $\Hom(0, 0)$, or the singleton set of the empty sequence:
  \begin{equation*}
    \begin{tikzcd}
      \epsilon
    \end{tikzcd}
  \end{equation*}
\end{example}

\begin{example}[$\Cube^1$]
  $\Cube^1$ consists of two points, given by $\Hom(0, 1)$, and a line, given by $\Hom(1, 1)$.
  \begin{equation*}
    \begin{tikzcd}
      L \arrow[r, "\kstar", dash] & R
    \end{tikzcd}
  \end{equation*}
\end{example}

>>>>

\documentclass{msc}

\usepackage{amsmath, amssymb, mathrsfs, wasysym, tikz, tikz-cd, lmodern, mathpazo, anyfontsize, xargs, environ, multirow, tabularx, caption, bookmark, booktabs, makecell, colortbl, minted, art.cls/colorpal, art.cls/ct, art.cls/sset, art.cls/lim, art.cls/joinargs}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[prefix=bonak]{art.cls/xkeymask}

@@

% Simplicial and Cubical sets
\newcommand{\DeltaHat}{\ensuremath{\hat{\boldsymbol{\Delta}}}}
\newcommand{\DeltaPlus}{\ensuremath{\boldsymbol{\Delta}_+}}
\newcommand{\Cube}{\ensuremath{\boldsymbol{{\square}}}}
\newcommand{\CSet}{\ensuremath{\mathsf{Set}_{\boldsymbol{\square}}}}

% Universe, referred to as U
\newcommandx{\U}[1][1=]{\ensuremath{\mathsf{\color{spanish-blue}{HSet}}_{#1}}}
\newcommandx{\HGpd}[1][1=]{\ensuremath{\mathsf{\color{spanish-blue}{HGpd}}_{#1}}}
\newcommand{\Type}{\ensuremath{\mathsf{\color{spanish-blue}{Type}}}}
\newcommand{\SProp}{\ensuremath{\mathsf{\color{spanish-blue}{SProp}}}}

% Examples
\newtheorem{example}[therm]{Example}

@@

\newcommand{\frametype}{\ensuremath{\mathsf{frametype}}}
\newcommand{\paintingtype}{\ensuremath{\mathsf{paintingtype}}}
\newcommand{\restrftype}{\ensuremath{\mathsf{restrftype}}}

\newcommand{\udensdash}[1]{%
    \tikz[baseline=(todotted.base)]{
        \node[inner sep=1pt,outer sep=0pt] (todotted) {$#1$};
        \draw[densely dashed] (todotted.south west) -- (todotted.south east);
    }%
}%

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\painting}[2][1,2]{\prim{painting}[][#1][#2][][]}
\newcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\newcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\newcommandx{\restrp}[4][1,2,3,4]{\restr{painting}[][#3][#4][#1][#2][][]}
\newcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\newcommandx{\cohl}{\coh{layer}[][][][][][][][]}
\newcommandx{\cohc}{\coh{painting}[][][][][][][][]}
\newcommandx{\coht}{\cohtwo{frame}[][][][][][][][]}
\newcommandx{\fullframe}[1][1]{\prim{fullframe}[][#1][][][]}

@@

% Caption setup
\DeclareCaptionFormat{plain}{#1#3}
\captionsetup{font=footnotesize,labelfont=bf}

% The eqntable environment
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\NewEnviron{eqntable}[1]{
  \fontsize{7.2}{9}\selectfont
  \begin{tabularx}{0.94\linewidth}{
    @{}
    >{$}l<{$}
    >{$}c<{$}
    >{$}c<{$}
    >{$}Y<{$}
    @{}}
    \toprule
    \BODY
    \bottomrule
  \end{tabularx}
  \captionof{table}{#1}
}

@@

% Double-spacing
% \linespread{2}

\begin{document}
\title{On the specific recursive structure of iterated parametricity}
\begin{authgrp}
  \author{Hugo Herbelin}
  \affiliation{Université Paris Cité, Inria, CNRS, IRIF, Paris
    \email{Hugo.Herbelin@inria.fr}}
  \author{Ramkumar Ramachandra}
  \affiliation{Unaffiliated \email{r@artagnon.com}}
\end{authgrp}
\begin{abstract}
  In a previous paper, we gave equations describing the construction of semi-simplicial and semi-cubical sets via iterated parametricity. The construction was formalised in the Rocq (formerly known as Coq) proof assistant building $n$-truncated sets using an induction on a large structure embedding equational reasoning.

  The current paper describes a new and lighter presentation of the construction where equational reasoning is replaced by definition. Being lighter, it can now be described in full details, closely following the corresponding new machine-checked formalisation.
\end{abstract}
\maketitle
\vspace{-2em} % Remove space after abstract

\section{Introduction}
A family of sets can commonly be represented in two ways: as a family properly speaking, indexed by the elements of a given set $S$, or as a set $T$ together with a map from $T$ to $S$, which specifies for each element of $T$ its dependency on $S$. In the former case, we call it an \emph{indexed} presentation. In the latter case, the set associated to a given element of $S$ is the fibre of this element, so we call it a \emph{fibred} presentation. The two presentations are equivalent and the equivalence can be phrased concisely in the language of homotopy type theory~\citep{hottbook} as the fibred/indexed equivalence\footnote{In an informal discussion, alternative nomenclatures were proposed: fibration/family equivalence and unbundled/bundled equivalence. The fibred/indexed nomenclature echoes the Grothendieck construction of fibred categories from indexed categories. The most elementary instance of the equivalence, with $\Type$ instead of $\U$, is sometimes called ``Grothendieck construction for dummies'', and its proof requires univalence~\citep{hottbook}.}.
\begin{equation*}
  \mbox{(fibred)}\qquad(\Sigma T: \U. (T \rightarrow S)) ~\simeq~ (S \rightarrow \U)\qquad \mbox{(indexed)}
\end{equation*}
Here, $\U$ represents in homotopy type theory the subset of types within a given universe where equality of any two elements has at most one proof.

A \emph{presheaf} on an category is a family of sets indexed by the object of the category with maps indexed by the morphisms. As such, it lives on the indexed side of the equivalence, contrasting with the fibred side, where we have \emph{discrete Grothendieck fibrations}~\citep{LoregianRiehl20}. However, there are situations where a presheaf can also be seen as living on the fibred side of the equivalence. This happens when the indexing category is \emph{direct}, or has a downwards-well-founded collection of non-identity morphisms. Let us consider, for instance, the case of a semi-cubical set~\citep{grandis03,buchholtz17} presented with $2n$ face maps from the set of $n$-cubes to the set of $(n-1)$-cubes. Formulated in type theory, the corresponding presheaf definition of a semi-cubical set prescribes a family of sets and face maps between them as follows.

@@

\section{The recursive structure of iterated parametricity}
Iterated parametricity (in indexed form) produces a family of type
families of the following form:
\begin{equation*}
  \begin{array}{llr}
    X_0 & :              & \U                                                                        \\
    X_1 & :              & X_0 \times X_0 \rightarrow  \U                                            \\
    X_2 & : \Pi a b c d. & X_1(a,b) \times X_1 (c,d) \times X_1(a,c) \times X_1 (b,d) \rightarrow \U \\
    \ldots
  \end{array}
\end{equation*}

Such family can be recursively described by the equations given on Table ?.

The well-foundedness of the construction can be justified by the observation that:
\begin{itemize}
  \item Defining $\framep[n][p]$ for $p \leq n$ requires the definition
        of $\painting[n-1][p-1]$, thus also $\framep[n-1][p-1]$ by typing
        dependency, as well as $\restrf[n][p-1]$, where
        $\restrf[n][p-1]$ in turn depends in $\framep[n][p-1]$ for its
        typing.

        Moreover, $\framep[n][p]$ depends on $\framep[n][p-1]$, so these
        dependencies are between the blocks of $\framep[n][k]$ for all
        $k\leq p$, as well as $\framep[n-1][k]$, $\painting[n-1][k]$ and
        $\restrf[n][k]$ for all $k\leq p-1$, which we respectively
        abbreviate as $\framep[n][0,...,p]$,
        $\framep[n-1][0,...,p-1]$, $\painting[n-1][0,...,p-1]$ and
        $\restrf[n][0,...,p-1]$.

        Also, $\restrf[n][p-1]$ depends on $n$.

        Schematically, using dotted arrows for dependencies in
        the type and plain arrows for dependencies in the definition, this
        gives the following dependencies for $p\leq n$ fixed:
        \begin{center}
          \begin{tikzcd}
            \framep[n][0,...,p] \arrow[d] \arrow[ddr]\\
            \painting[n-1][0,...,p-1] \arrow[d, dotted] \\
            \framep[n-1][0,...,p-1] &
            \restrf[n][0,...,p-1] \arrow[l, dotted] \arrow[uul, dotted, "\mbox{$[0,...,p-1]$}"', near start, shift right=2mm, hook']\\
          \end{tikzcd}
        \end{center}

        In particular, the mutual dependency on one side of
        $\framep[n][0,...,p-1]$ within the type of
        $\restrf[n][0,...,p-1]$ and on the other side of
        $\restrf[n][0,...,p-1]$ within the definition of
        $\framep[n][0,...,p]$ requires to mutually define the type of
        $\restrf[n][0,...,p-1]$ together with $\framep[n][p]$ (or
        $\framep[n][0,...,p]$) as a function from
        $\restrf[n][0,...,p-1]$.

        To talk about sequences of $\framep$, $\painting$ or $\restrf$,
        we need to specify their types. The type of $\restrf$ will be
        defined mutually with $\framep$ of the same level, but the type of
        $\framep$, $\painting$ can be defined in advance, the second one
        being however dependent on the first one. In general, components of
        such sequences will have a type dependent on the previous components
        of the sequence, as it is the case for $\restrf$. So, the type of
        a sequence of length $p$ will be expressed as a $n$-iterated
        $\Sigma$-type. In the case of $\framep$ and $\painting$, there will
        be however no dependency, so an ordinary product can be used. If
        $\framep[n][0,...,p]$ is such a sequence of length $p+1$, we write
        $\framep[n][0,...,p-1]$ for its immediate subsequence of length $p$
        and $\framep[n][p]$ for its last component. We can now define the
        types of sequences of $\framep$ and $\painting$ as follows:

        $$
          \begin{array}{llcl}
            \frametype^{n,[0,...,p]}  &                                                     & :      & Type_{l+1}                                                               \\
            \frametype^{n,[]}         &                                                     & \defeq & unit                                                                     \\
            \frametype^{n,[0,...,p]}  &                                                     & \defeq & \frametype^{n,[0,...,p-1]} \times Type_l                               \\
            \\
            \paintingtype^{n,[0,...,p]} & (\framep[n][0,...,p]:\frametype^{n,[0,...,p]}) & :      & Type_{l+1}                                                               \\
            \paintingtype^{n,[]}        & \star                                               & \defeq & unit                                                                     \\
            \paintingtype^{n,[0,...,p]} & (\framep[n][0,...,p])                            & \defeq & \paintingtype^{n,[0,...,p-1]}(\framep[n][0,...,p-1]) \times (\framep[n][p] \rightarrow Type_l) \\
          \end{array}
        $$

        Formally, this can be described as:
        $$
          \begin{array}{llcl}
            \framep[n][0,...,p]                                     &
            \multicolumn{3}{l}{
              \left(\begin{array}{l}
                        \framep[n-1][0,...,p-1]:\frametype^{n-1,[0,...,p-1]}                               \\
                        \painting[n-1][0,...,p-1]:\paintingtype^{n-1,[0,...,p-1]}(\framep[n-1][0,...,p-1]) \\
                        \restrf[n][0,...,p-1]:\restrftype^{n,[0,...,p-1]}
                                \left(\begin{array}{l}
                                    \framep[n-1][0,...,p-1]  \\
                                    \painting[n-1][0,...,p-1] \\
                                  \end{array}\right)                             \\
                      \end{array}\right) : \frametype^{n-1,[0,...,p]}
            }                                                                                         \\
            \framep[n][0]                                           &
            \left(\begin{array}{l}
                      \star \\
                      \star \\
                      \star \\
                    \end{array}\right)                                     & \defeq & (\star,unit)      \\
            \framep[n][0,...,p+1]                                   &
            \left(\begin{array}{l}
                      \framep[n-1][0,...,p]  \\
                      \painting[n-1][0,...,p] \\
                      \restrf[n][0,...,p] \\
                    \end{array}\right) & \defeq &
            \left(\begin{array}{l}
                      \framep[n][0,...,p]
                      \left(\begin{array}{l}
                          \framep[n-1][0,...,p-1]  \\
                          \painting[n-1][0,...,p-1] \\
                          \restrf[n][0,...,p-1] \\
                        \end{array}\right)     , \\
                      \Sigma d:\framep[n][0,...,p]
                      \left(\begin{array}{l}
                          \framep[n-1][0,...,p-1]  \\
                          \painting[n-1][0,...,p-1] \\
                          \restrf[n][0,...,p-1] \\
                        \end{array}\right).\, \\
                      \quad \Pi\omega.\,\painting[n-1][p](\restrf[n][p](d))     \\
                    \end{array}\right)                    \\
            \restrftype^{n,[0,...,p-1]}                            &
            \multicolumn{3}{l}{
              \left(\begin{array}{l}
                        \framep[n-1][0,...,p-1]:\frametype^{n-1,[0,...,p-1]}                               \\
                        \painting[n-1][0,...,p-1]:\paintingtype^{n-1,[0,...,p-1]}(\framep[n-1][0,...,p-1]) \\
                      \end{array}\right) : Type_{l}
            }                                                                                         \\
            \restrftype^{n,[]}                                     &
            \left(\begin{array}{l}
                      \star \\
                      \star \\
                    \end{array}\right)                                     & \defeq &
            unit                                                                                      \\
            \restrftype^{n,[0,...,p]}                              &
            \left(\begin{array}{l}
                      \framep[n-1][0,...,p]  \\
                      \painting[n-1][0,...,p] \\
                    \end{array}\right) & \defeq &
            \begin{array}{l}
              \Sigma \restrf[n][0,...,p-1]                            \\
              \quad:\restrftype^{n,[0,...,p-1]}
              \left(\begin{array}{l}
                        \framep[n-1][0,...,p-1]  \\
                        \painting[n-1][0,...,p-1] \\
                      \end{array}\right)
              .\,                                                           \\
              \Pi q\leq n.\,\Pi \omega.\,\framep[n][0,...,p]
              \left(\begin{array}{l}
                        \framep[n-1][0,...,p-1]  \\
                        \painting[n-1][0,...,p-1] \\
                        \restrf[n][0,...,p-1] \\
                      \end{array}\right) \\
            \end{array} \\
          \end{array}
        $$
          \item Defining $\painting[n][p]$, in addition to the type dependency
        in $\framep[n][p]$, requires the definition of
        $\painting[n-1][p-1]$, thus also $\framep[n-1][p-1]$ by typing
        dependency, as well as $\restrf[n][p-1]$. Since
        $\painting[n][p]$ also depends on $\painting[n][p+1]$ when $p<n$,
        these dependencies are actually between $\painting[n][p,...,n]$,
        $\painting[n-1][p,...,n-1]$, $\framep[n][p,...,n]$,
        $\framep[n-1][p,...,n-1]$ and $\restrf[n][p,...,n]$. Also,
        for the case $p=n$, this also requires the assumption of an
        inhabitant $X_{n}$ of $\nu$Set$^{=n}$. Finally,
        $\framep[n][p,...,n]$, $\framep[n-1][p,...,n-1]$ and
        $\restrf[n][p,...,n-1]$ respectively depend on the earlier
        values $\framep[n][0,...,p-1]$, $\framep[n-1][0,...,p-1]$ and
        $\restrf[n][0,...,p-1]$.  Below, we use dashed underline and
        dashed arrows for dependencies in subconstructions, here in the
        construction of $\framep$ at level $n-1$ and $\restrftype$ at
        level $n$:
        \begin{center}
          \begin{tikzcd}
            & & \painting[n][p,...,n] \arrow[dd, bend right=70] \arrow[dddr] \arrow[d, dotted]\\
            \udensdash{\framep}^{n,[0,...,p-1]} \arrow[d, dashed] \arrow[ddr, dashed] & &
            \udensdash{\framep}^{n,[p,...,n]} \arrow[ll, dotted] \arrow[d, dashed] \arrow[ddr, dashed, shift right=2mm] \\
            \painting[n-1][0,...,p-1] \arrow[d, dotted] & &
            \painting[n-1][p,...,n-1] \arrow[ll, dotted] \arrow[d, dotted] \\
            \framep[n-1][0,...,p-1] &
            \udensdash{\restrf}^{n,[0,...,p-1]} \arrow[l, dotted] \arrow[uul, dotted, shift right=2mm, hook'] &
            \framep[n-1][p,...,n-1] \arrow[ll, dotted, bend left=15] &
            \udensdash{\restrf}^{n,[p,...,n-1]},X_n \arrow[ll, dotted, bend left=15] \arrow[l, dotted] \arrow[uul, dotted, hook']\\
          \end{tikzcd}
        \end{center}

To define $\painting$ formally, we need to define $\framep[n][p,...,n]$, $\painting[n][p,...,n]$ and $\restrf[n][p,...,n]$:

        $$
          \begin{array}{llcl}
            \frametype^{n,[p,...,n]}  &                                                     & :      & Type_{l+1}                                                               \\
            \frametype^{n,[]}         &                                                     & \defeq & unit                                                                     \\
            \frametype^{n,[p,...,n]}  &                                                     & \defeq & Type_l \times \frametype^{n,[p+1,...,n-1]}                               \\
            \\
            \paintingtype^{n,[p,...,n]} & (\framep[n][p,...,n]:\frametype^{n,[p,...,n]}) & :      & Type_{l+1}                                                               \\
            \paintingtype^{n,[]}        & \star                                               & \defeq & unit                                                                     \\
            \paintingtype^{n,[p,...,n]}_{p<n} & (\framep[n][p,...,n])                            & \defeq & (\framep[n][p] \rightarrow Type_l) \times \paintingtype^{n,[p+1,...,n]}(\framep[n][p+1,...,n]) \\
            \\
            \restrftype^{n,[p,...,n]} &
            \multicolumn{3}{l}{
              \left(\begin{array}{l}
                        \framep[n-1][0,...,p-1]:\frametype^{n-1,[0,...,p-1]}                               \\
                        \painting[n-1][0,...,p-1]:\paintingtype^{n-1,[0,...,p-1]}(\framep[n-1][0,...,p-1]) \\
                        \framep[n-1][p,...,n-1]:\frametype^{n-1,[p,...,n-1]}                                     \\
                        \painting[n-1][p,...,n-1]:\paintingtype^{n-1,[p,...,n-1]}(\framep[n-1][p,...,n-1])           \\
                        \restrf[n][0,...,p-1]:\restrftype^{n,[0,...,p-1]}
